---
title: "01 Machine Learning Fundamentals"
author: "Samim Abbasi"
date: "2021-01-03"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    collapsed: false
    code_folding: hide
    number_sections: true
    toc_depth: 3
    #code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    message = FALSE,
    warning = FALSE
    )
```

# Load libraries
```{r}
# install.packages("plotly")

library(tidyverse)
#library(ggplot2)
#library(dplyr)
#library(forcats)
#library(tibble)
#library(stringr)
#library(purrr)
library(readr)
library(tidyr)
library(tidyquant)
library(broom)
library(umap)
library(ggrepel)
library(dplyr)
library(magrittr)
library(cluster) 
```


# Challenge Summary

__Your organization wants to know which companies are similar to each other to help in identifying potential customers of a SAAS software solution (e.g. Salesforce CRM or equivalent) in various segments of the market. The Sales Department is very interested in this analysis, which will help them more easily penetrate various market segments.__

You will be using stock prices in this analysis. You come up with a method to classify companies based on how their stocks trade using their daily stock returns (percentage movement from one day to the next). This analysis will help your organization determine which companies are related to each other (competitors and have similar attributes). 

You can analyze the stock prices using what you've learned in the unsupervised learning tools including K-Means and UMAP. You will use a combination of `kmeans()` to find groups and `umap()` to visualize similarity of daily stock returns.




# Data

We will be using stock prices in this analysis. Although some of you know already how to use an API to retrieve stock prices I obtained the stock prices for every stock in the S&P 500 index for you already. The files are saved in the `session_6_data` directory. 

We can read in the stock prices. The data is 1.2M observations. The most important columns for our analysis are:

- `symbol`: The stock ticker symbol that corresponds to a company's stock price
- `date`: The timestamp relating the symbol to the share price at that point in time
- `adjusted`: The stock price, adjusted for any splits and dividends (we use this when analyzing stock data over long periods of time) 


```{r}
# STOCK PRICES
sp_500_prices_tbl <- read_rds("/home/abbasi/Desktop/R/BDML/sp_500_prices_tbl.rds")
sp_500_prices_tbl
sp_500_prices_tbl <- as_tibble(sp_500_prices_tbl)
```

The second data frame contains information about the stocks the most important of which are:

- `company`: The company name
- `sector`: The sector that the company belongs to

```{r}
# SECTOR INFORMATION
sp_500_index_tbl <- read_rds("/home/abbasi/Desktop/R/raw_data/sp_500_index_tbl.rds")
sp_500_index_tbl
sp_500_index_tbl <- as_tibble(sp_500_index_tbl)
sp_500_index_tbl
```


# Question

<mark>Which stock prices behave similarly?</mark>

Answering this question helps us __understand which companies are related__, and we can use clustering to help us answer it!

Even if you're not interested in finance, this is still a great analysis because it will tell you which companies are competitors and which are likely in the same space (often called sectors) and can be categorized together. Bottom line - This analysis can help you better understand the dynamics of the market and competition, which is useful for all types of analyses from finance to sales to marketing.  

Let's get started. 

# Objectives

Apply your knowledge on K-Means and UMAP along with `dplyr`, `ggplot2`, and `purrr` to create a visualization that identifies subgroups in the S&P 500 Index. You will specifically apply:

- Modeling: `kmeans()` and `umap()`
- Iteration: `purrr`
- Data Manipulation: `dplyr`, `tidyr`, and `tibble`
- Visualization: `ggplot2` (bonus `plotly`)

## Step 1 - Convert stock prices to a standardized format (daily returns)

What you first need to do is get the data in a format that can be converted to a "user-item" style matrix. The challenge here is to connect the dots between what we have and what we need to do to format it properly.

We know that in order to compare the data, it needs to be standardized or normalized. Why? Because we cannot compare values (stock prices) that are of completely different magnitudes. In order to standardize, we will convert from adjusted stock price (dollar value) to daily returns (percent change from previous day). Here is the formula. 

$$ 
return_{daily} = \frac{price_{i}-price_{i-1}}{price_{i-1}}
$$

First, what do we have? We have stock prices for every stock in the [SP 500 Index](https://finance.yahoo.com/quote/%5EGSPC?p=%5EGSPC), which is the daily stock prices for over 500 stocks. The data set is over 1.2M observations. 

```{r}
sp_500_prices_tbl %>% glimpse()

```{r}
# STOCK PRICES
sp_500_prices_tbl <- read_rds("/home/abbasi/Desktop/R/BDML/sp_500_prices_tbl.rds")
  
  
  

```

- Select the `symbol`, `date` and `adjusted` columns
- Filter to dates beginning in the year 2018 and beyond. 
- Compute a Lag of 1 day on the adjusted stock price. Be sure to group by symbol first, otherwise we will have lags computed using values from the previous stock in the data frame. 
- Remove a `NA` values from the lagging operation
- Compute the difference between adjusted and the lag
- Compute the percentage difference by dividing the difference by that lag. Name this column `pct_return`.
- Return only the `symbol`, `date`, and `pct_return` columns
- Save as a variable named `sp_500_daily_returns_tbl`


```{r}
## Applying data transformation

sp_500_daily_tbl <- sp_500_prices_tbl %>% 
  
  select(symbol, date, adjusted) %>%
  filter(date >= "2018-01-01"  ) %>%
  group_by(symbol) %>%
  mutate(ladjusted = lag(adjusted)) %>%
  filter(!is.na(ladjusted)) %>%
  mutate(differenc = adjusted - ladjusted) %>%
  mutate(pct_return = differenc/ladjusted) %>%
  select(symbol, date, pct_return) %>%
  glimpse()



```
## Step 2 - Convert to User-Item Format

Now that we have the daily returns (percentage change from one day to the next), we can convert to a user-item format. The user in this case is the `symbol` (company), and the item in this case is the `pct_return` at each `date`. 

- Spread the `date` column to get the values as percentage returns. Make sure to fill an `NA` values with zeros. 
- Save the result as `stock_date_matrix_tbl`


```{r}
stock_date_matrix_tbl <- sp_500_daily_tbl %>% 
  
  pivot_wider(names_from = date, values_from = pct_return, values_fill = 0 ) %>%
  ungroup()


  
```
## Step 3 - Perform K-Means Clustering

```{r}
kmeans_obj <- stock_date_matrix_tbl %>% 
  select(-symbol) %>% 
  filter_all(all_vars(!is.infinite(.))) %>%
  filter_all(all_vars(!is.na(.))) %>%
  
  kmeans(centers =  4 ,nstart = 20) 
#kmeans_obj$cluster
```

```{r}
#broom::tidy(kmeans_obj) %>% glimpse()
```
```{r}
broom::glance(kmeans_obj)

```


## Step 4 - Find the optimal value of K


Now that we are familiar with the process for calculating `kmeans()`, let's use `purrr` to iterate over many values of "k" using the `centers` argument. 

```{r}
# Use purrr to map
kmeans_mapper <- function(centers = 3) {
    
        stock_date_matrix_tbl %>%
           select(-symbol) %>%
           filter_all(all_vars(!is.infinite(.))) %>%
           filter_all(all_vars(!is.na(.))) %>% 
           kmeans(centers = centers, nstart = 100) 
          
        
}
3 %>% kmeans_mapper() %>% glance()

```


```{r}
# Output: k_means_mapped_tbl 
k_means_mapped_tbl <- tibble(centers = 1:30) %>%
    mutate(k_means = centers %>% map(kmeans_mapper)) %>%
    mutate(glance  = k_means %>% map(glance))
```

```{r}
k_means_mapped_tbl %>%
    unnest(glance) %>%
    select(centers, tot.withinss)
```
```{r}
# Visualize Scree Plot


k_means_mapped_tbl %>%
    unnest(glance) %>%
    select(centers, tot.withinss) %>%
    
    # Visualization
    ggplot(aes(centers, tot.withinss)) +
    geom_point(color = "#2DC6D6", size = 4) +
    geom_line(color = "#2DC6D6", size = 1) +
    # Add labels (which are repelled a little)
    ggrepel::geom_label_repel(aes(label = centers), color = "#2DC6D6") + 
    
    # Formatting
    labs(title = "Skree Plot",
    subtitle = "Measures the distance each of the stock are from the closest K-Means center",
    caption = "Conclusion: Based on the Scree Plot, we select 4 clusters to segment the customer base.")

```



We can see that the Scree Plot becomes linear (constant rate of change) between 5 and 10 centers for K.


## Step 5 - Apply UMAP


Next, let's plot the `UMAP` 2D visualization to help us investigate cluster assignments. 

```{r}
umap_results <- stock_date_matrix_tbl %>% 
  select(-symbol) %>% 
  filter_all(all_vars(!is.infinite(.))) %>%
  filter_all(all_vars(!is.na(.))) %>%
  umap() 

```

```{r}
stock_date_matrix_tbl1 <- stock_date_matrix_tbl %>% 
  filter_all(all_vars(!is.infinite(.))) %>%
  filter_all(all_vars(!is.na(.))) 
```


```{r}
# Convert umap results to tibble with symbols
umap_results_tbl <- umap_results$layout %>%
    as_tibble(.name_repair = "unique") %>% 
    set_names(c("x", "y")) %>%
    bind_cols(
        stock_date_matrix_tbl1 %>% select(symbol) 
    )

umap_results_tbl %>%
    ggplot(aes(x, y)) +
    geom_point() + 
    geom_label_repel(aes(label = symbol), size = 5)



# Output: umap_results_tbl

```

We can now see that we have some clusters. However, we still need to combine the K-Means clusters and the UMAP 2D representation.




## Step 6 - Combine K-Means and UMAP

Next, we combine the K-Means clusters and the UMAP 2D representation


First, pull out the K-Means for 10 Centers. Use this since beyond this value the Scree Plot flattens. 

```{r}
# Get the k_means_obj from the 10th center

k_means_obj <- k_means_mapped_tbl %>%
    pull(k_means) %>%
    pluck(10)
# Store as k_means_obj
```

Next, we'll combine the clusters from the `k_means_obj` with the `umap_results_tbl`.

- Begin with the `k_means_obj`
- Augment the `k_means_obj` with the `stock_date_matrix_tbl` to get the clusters added to the end of the tibble
- Select just the `symbol` and `.cluster` columns
- Left join the result with the `umap_results_tbl` by the `symbol` column
- Left join the result with the result of `sp_500_index_tbl %>% select(symbol, company, sector)` by the `symbol` column. 
- Store the output as `umap_kmeans_results_tbl`

```{r}

umap_kmeans_results_tbl_i <- k_means_obj %>% 
    # Augment the k_means_obj with the stock_date_matrix_tbl to get the clusters added to the end of the tibble
    augment(stock_date_matrix_tbl1) %>%
    # Select just the symbol and .cluster columns
    select(symbol, .cluster)  

```

```{r}
sp_500_index_tbl <- read_rds("/home/abbasi/Desktop/R/raw_data/sp_500_index_tbl.rds")
  # Left join the result with the umap_results_tbl by the symbol 
umap_kmeans_results_tbl1  <- left_join(umap_kmeans_results_tbl_i, umap_results_tbl, by = "symbol")

  #Left join the result with the result of sp_500_index_tbl %>% select(symbol, company, sector) by the symbol column. 
subsub_sp_500_prices_tbl <- sp_500_index_tbl %>% select(symbol, company, sector) 

umap_kmeans_results_tbl2  <- left_join(umap_kmeans_results_tbl1,subsub_sp_500_prices_tbl, by = "symbol") 
umap_kmeans_results_tbl <- umap_kmeans_results_tbl2    

# Output: umap_kmeans_results_tbl 
```

```{r}

umap_kmeans_results_tbl %>%
    mutate(label_text = str_glue("stock: {symbol}
                                 Cluster: {.cluster}")) %>%
    
    ggplot(aes(x, y, color = .cluster)) +
    
    # Geometries
    geom_point() +
    geom_label_repel(aes(label = label_text), size = 2, fill = "#282A36") +
    
    # Formatting
    scale_color_manual(values=c("red", "blue", "green", "black", "yellow", "orange", "cyan", "magenta", "plum", "pink")) +
    
    labs(title = "Stock Segmentation: 2D Projection",
    subtitle = "UMAP 2D Projection with K-Means Cluster Assignment",
    caption = "Conclusion: 10 Customer Segments identified using 2 algorithms") +
    theme(legend.position = "none")
```









