---
title: "03 Automated Machine Learning with H20 I"
date: "2021-12-28"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    collapsed: false
    number_sections: true
    toc_depth: 3
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```




# Automated Machine Learning with H20 (I)

1. Compensation Features
What can you deduce about the interaction between Monthly Income and Attrition?
c. Those that are leaving have a lower Monthly Income

2. Compensation Features
What can you deduce about the interaction between Percent Salary Hike and Attrition?
c. Those that are leaving have lower Percent Salary Hike

3. Compensation Features
What can you deduce about the interaction between Stock Option Level and Attrition?
c. It's difficult to deduce anything based on the visualization

4. Survey Results
What can you deduce about the interaction between Environment Satisfaction and Attrition?
b. A higher proportion of those leaving have a high environment satisfaction level

5. Survey Results
What can you deduce about the interaction between Work Life Balance and Attrition
a. Those that are leaving have higher density of 2's and 3's

6. Performance Data
What Can you deduce about the interaction between Job Involvement and Attrition?
b. Those that are leaving have a lower density of 1's and 2's

7. Work-Life Features
What can you deduce about the interaction between Over Time and Attrition?
b. The proportion of those staying that are working Over Time are high compared to those that are not staying

8. Training and Education
What can you deduce about the interaction between Training Times Last Year and Attrition
b. People that leave tend to have less annual training

9. Time-Based Features
What can you deduce about the interaction between Years At Company and Attrition
b. People that leave tend to have less working years at the company


10. Time-Based Features
What can you deduce about the interaction between Years Since Last Promotion and Attrition?
b. Those that are leaving have fewer years since last promotion than those that are staying


# Automated Machine Learning with H20 (II)



```{r}
library(modeldata)
library(readr)
library(readxl)
library(modelr)
library(modeltools)
library(tidymodels)
library(magrittr)
library(dplyr)
library(sjmisc)
library(magrittr)
library(haven)
library(sjlabelled)
library(rsample)
library(recipes)
library(rstanarm)
library(broom.mixed)
library(h2o)
library(bayesplot)
h2o.init()
```


```{r}
theme_set(bayesplot::theme_default())
product_backorders_tbl <- read_csv("/home/abbasi/Desktop/R/raw_data/product_backorders.csv") 
product_backorders_tbl %>% glimpse()
data_split <- initial_split(product_backorders_tbl, prop = 3/4)
# Assign training and test data
train_data <- training(data_split)
test_data  <- testing(data_split)
factor_names <- c("went_on_backorder")
product_rec <- 
  recipe(went_on_backorder ~ ., data = train_data) %>%  
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_predictors()) %>% 
  step_mutate_at(went_on_backorder, fn = as.factor) %>%
  prep()
d <- summary(product_rec)
train_tbl <- bake(product_rec, new_data = train_data)
test_tbl  <- bake(product_rec, new_data = test_data)
#train_tbl <- train(product_rec, new_data = train_data)
#test_tbl  <- train(product_rec, new_data = test_data)
```

# Modeling
```{r}
h2o.init()
# Split data into a training and a validation data frame
# Setting the seed is just for reproducability
split_h2o <- h2o.splitFrame(as.h2o(train_tbl), ratios = c(0.75), seed = 1234)
train_h2o <- split_h2o[[1]]
valid_h2o <- split_h2o[[2]]
test_h2o  <- as.h2o(test_tbl)
# Set the target and predictors
y <- "went_on_backorder"
x <- setdiff(names(train_h2o), y)
```

```{r}
automl_models_h2o <- h2o.automl(
  x = x,
  y = y,
  training_frame    = train_h2o,
  validation_frame  = valid_h2o,
  leaderboard_frame = test_h2o,
  max_runtime_secs  = 30,
  nfolds            = 5 
)
```

```{r}
typeof(automl_models_h2o)
slotNames(automl_models_h2o)
automl_models_h2o@leaderboard 
automl_models_h2o@leader 
```

```{r}
h2o.init()
 h2o.getModel("StackedEnsemble_AllModels_AutoML_20210111_194353") %>%
h2o.saveModel(path = "/home/abbasi/Desktop/R/h20_models/")
 h2o.getModel("XGBoost_grid__1_AutoML_20210111_194353_model_1") %>%
h2o.saveModel(path = "/home/abbasi/Desktop/R/h20_models/")
  h2o.getModel("StackedEnsemble_BestOfFamily_AutoML_20210111_194353") %>%
h2o.saveModel(path = "/home/abbasi/Desktop/R/h20_models/")
   h2o.getModel("XGBoost_grid__1_AutoML_20210111_194353_model_2") %>%
h2o.saveModel(path = "/home/abbasi/Desktop/R/h20_models/")
    h2o.getModel("XGBoost_3_AutoML_20210111_194353") %>%
h2o.saveModel(path = "/home/abbasi/Desktop/R/h20_models/")

```


```{r}
extract_h2o_model_name_by_position <- function(h2o_leaderboard, n = 1, verbose = T) {
  
  model_name <- h2o_leaderboard %>%
    as_tibble() %>%
    slice(n) %>%
    pull(model_id)
  
  if (verbose) message(model_name)
  
  return(model_name)
  
}
```




```{r}
h2o.init()
stacked_ensemble_h2o <- h2o.loadModel("/home/abbasi/Desktop/R/h20_models/XGBoost_3_AutoML_20210111_194353")
stacked_ensemble_h2o
predictions <- h2o.predict(stacked_ensemble_h2o, newdata = as.h2o(test_tbl))
typeof(predictions)
predictions_tbl <- predictions %>% as_tibble()
predictions_tbl
```
